---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# jentre

<!-- badges: start -->
<!-- badges: end -->

jentre is a client for the NCBI's Entrez API.

The Entrez API has many quirks. jentre attempts to deal with those while presenting
a convenient interface. It is designed for bulk metadata fetching and link traversal,
though also provides full access to other parts of Entrez, albeit with fewer helpers.

Features of jentre:

  * Provides objects representing sets of Entrez identifiers to avoid mixing them up
  * Batches requests behind the scenes when needed
  * Based on {httr2} and {xml2}

The {rentrez} package is more mature and might suit a broader set of applications.


## Installation

You can install the development version of jentre like so:

``` r
# development version
install.packages('jentre', repos = c('https://cidm-ph.r-universe.dev', 'https://cloud.r-project.org'))
```

## Example

```{r example}
library(jentre)

# if needed, you can construct an arbitrary request
req <- entrez_request("efetch.fcgi", id = "")
req
# you'll need to execute it yourself:
# httr2::req_perform(req)
```
